An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network

### 2025-07-16 paper/main.tex
layout

\begin{abstract}
	We study how the numerical precision of weights, activations, and gradients in neural networks can be reduced without sacrificing too much of the classification performance. Using MNIST as a dataset, we benchmark uniform 64-16-bit quantisation, mixed-precision formats, and post-training decimal rounding. The results reveal that (i) 16-bit end-to-end training preserves full accuracy for the MNIST dataset and the network used, (ii) three-decimal-place weight rounding while keeping 32-bit precision incurs minimal accuracy loss. These findings verify limits for memory- and energy-constrained usecases like inference in mobile devices.
\end{abstract}

@@

@@

\section{Results}
In the first experiment, every floating-point tensor that is used inside the
training loop was converted to a lower precision data type. This therefore
includes the weights and biases, activations, gradients, optimizer states, etc.
The types tried were float64, float32, and float16.

>>>>

\begin{abstract}
	We study how the numerical precision of weights, activations, and gradients in neural
	networks can be reduced without sacrificing too much of the classification
	performance. Using MNIST as a dataset, we benchmark uniform 64-16-bit quantisation,
	mixed-precision formats, and post-training decimal rounding. The results reveal that
	(i) 16-bit end-to-end training does not significantly decrease test accuracy for the
	MNIST dataset and the network used, (ii) three-decimal-place weight rounding while
	keeping 32-bit precision incurs minimal accuracy loss. These findings verify limits
	for memory- and energy-constrained usecases like inference in mobile devices.
\end{abstract}

@@

\newpage

@@

\newpage
\section{Results}
In the first experiment, every floating-point tensor that is used inside the
training loop was converted to a lower precision data type. This therefore
includes the weights and biases, activations, gradients, optimizer states, etc.
The types tried were float64, float32, and float16.
